% arara: xelatex
\documentclass[12pt]{article}

% \usepackage{physics} % это трэшовый пакет!

\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
    pdftitle={Overleaf Example},
    pdfpagemode=FullScreen,
    }
\urlstyle{same} % шрифт из основного текста

\usepackage{verse}

\usepackage{ulem} % перечёркнутый шрифт \sout{что-то} 

\usepackage{tikzducks}

\usepackage{tikz} % картинки в tikz
\usetikzlibrary{arrows.meta}
\usetikzlibrary{shapes, arrows, positioning}
\usepackage{tikz-3dplot} 
\usepackage{tkz-euclide}

\usepackage{verbatim}
\usepackage{microtype} % свешивание пунктуации

\usepackage{array} % для столбцов фиксированной ширины

\usepackage{indentfirst} % отступ в первом параграфе

\usepackage{sectsty} % для центрирования названий частей
\allsectionsfont{\centering}

\usepackage{amsmath, amsfonts, amssymb, amsthm} % куча стандартных математических плюшек

\usepackage{comment}

\usepackage[top=2cm, left=1.2cm, right=1.2cm, bottom=2cm]{geometry} % размер текста на странице

\usepackage{lastpage} % чтобы узнать номер последней страницы

\usepackage{enumitem} % дополнительные плюшки для списков
%  например \begin{enumerate}[resume] позволяет продолжить нумерацию в новом списке
\usepackage{caption}

\usepackage{url} % to use \url{link to web}

\newcommand{\smallduck}{\begin{tikzpicture}[scale=0.3]
    \duck[
        cape=black,
        hat=black,
        mask=black
    ]
    \end{tikzpicture}}

\usepackage{fancyhdr} % весёлые колонтитулы
\pagestyle{fancy}
\lhead{}
\chead{}
\rhead{}
\lfoot{}
\cfoot{}
\rfoot{}

\renewcommand{\headrulewidth}{0.4pt}
\renewcommand{\footrulewidth}{0.4pt}

\usepackage{tcolorbox} % рамочки!
\usepackage{graphicx} % рисунки
\usepackage{todonotes} % для вставки в документ заметок о том, что осталось сделать
% \todo[inline]{Здесь надо коэффициенты исправить}
% \missingfigure{Здесь будет Последний день Помпеи}
% \listoftodos - печатает все поставленные \todo'шки


% более красивые таблицы
\usepackage{booktabs}
% заповеди из докупентации:
% 1. Не используйте вертикальные линни
% 2. Не используйте двойные линии
% 3. Единицы измерения - в шапку таблицы
% 4. Не сокращайте .1 вместо 0.1
% 5. Повторяющееся значение повторяйте, а не говорите "то же"


\setcounter{MaxMatrixCols}{20}
% by crazy default pmatrix supports only 10 cols :)


\usepackage{fontspec}
\usepackage{libertine}
\usepackage{polyglossia}

\setmainlanguage{english}
\setotherlanguages{russian}

% download "Linux Libertine" fonts:
% http://www.linuxlibertine.org/index.php?id=91&L=1
% \setmainfont{Linux Libertine O} % or Helvetica, Arial, Cambria
% why do we need \newfontfamily:
% http://tex.stackexchange.com/questions/91507/
% \newfontfamily{\cyrillicfonttt}{Linux Libertine O}

%\AddEnumerateCounter{\asbuk}{\russian@alph}{щ} % для списков с русскими буквами
%\setlist[enumerate, 1]{label=\asbuk*),ref=\asbuk*}
% \setlist[enumerate, 1]{label=\alph*),ref=\alph*}



\usepackage[bibencoding=auto, backend=biber, sorting=none, style=alphabetic]{biblatex}
\addbibresource{cgt.bib}

\graphicspath{{./figures}}

%% эконометрические сокращения
\DeclareMathOperator{\Cov}{\mathbb{C}ov}
\DeclareMathOperator{\Corr}{\mathbb{C}orr}
\DeclareMathOperator{\sCov}{sCov}
\DeclareMathOperator{\sCorr}{sCorr}
\DeclareMathOperator{\sVar}{sVar}
\DeclareMathOperator{\Var}{\mathbb{V}ar}
\DeclareMathOperator{\hVar}{\widehat{\Var}}
\DeclareMathOperator{\hCov}{\widehat{\Cov}}
\DeclareMathOperator{\hCorr}{\widehat{\Corr}}
\DeclareMathOperator{\col}{col}
\DeclareMathOperator{\colspan}{colspan}

\DeclareMathOperator{\se}{se}
\DeclareMathOperator{\row}{row}
\DeclareMathOperator{\grad}{grad}
\DeclareMathOperator{\BestLin}{BestLin}

\let\P\relax
\DeclareMathOperator{\P}{\mathbb{P}}

\DeclareMathOperator{\E}{\mathbb{E}}
\DeclareMathOperator{\trace}{trace}
\DeclareMathOperator{\rank}{rank}
\DeclareMathOperator{\tr}{\trace}
\DeclareMathOperator{\rk}{\rank}
\DeclareMathOperator{\card}{card}
\DeclareMathOperator{\Span}{span}
% \span используется латехом для внутренних нужд

\DeclareMathOperator{\mgf}{mgf}

\DeclareMathOperator{\Convex}{Convex}
\DeclareMathOperator{\plim}{plim}

\usepackage{mathtools}
\DeclarePairedDelimiter{\norm}{\lVert}{\rVert}
\DeclarePairedDelimiter{\abs}{\lvert}{\rvert}
\DeclarePairedDelimiter{\scalp}{\langle}{\rangle}
\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}

% https://tex.stackexchange.com/questions/12910/how-do-i-typeset-vertical-and-horizontal-lines-inside-a-matrix
% для линий в матрицах
\newcommand*{\horzbar}{\rule[.5ex]{2.5ex}{0.5pt}}
\newcommand*{\vertbar}{\rule[-1ex]{0.5pt}{2.5ex}}


\newcommand{\cN}{\mathcal{N}}
\newcommand{\cF}{\mathcal{F}}

\newcommand{\RR}{\mathbb{R}}
\newcommand{\NN}{\mathbb{N}}

\newcommand{\dBern}{\mathrm{Bern}}
\newcommand{\dPois}{\mathrm{Pois}}
\newcommand{\dBin}{\mathrm{Bin}}
\newcommand{\dMult}{\mathrm{Mult}}
\newcommand{\dGeom}{\mathrm{Geom}}
\newcommand{\dNHGeom}{\mathrm{NHGeom}}
\newcommand{\dHGeom}{\mathrm{HGeom}}
\newcommand{\dDUnif}{\mathrm{DUnif}}
\newcommand{\dFS}{\mathrm{FS}}
\newcommand{\dNBin}{\mathrm{NBin}}

\newcommand{\dTri}{\mathrm{Triangle}}
\newcommand{\dUnif}{\mathrm{Unif}}
\newcommand{\dCauchy}{\mathrm{Cauchy}}
\newcommand{\dN}{\mathcal{N}}
\newcommand{\dLN}{\mathcal{LN}}
\newcommand{\dExpo}{\mathrm{Expo}}
\newcommand{\dBeta}{\mathrm{Beta}}
\newcommand{\dGamma}{\mathrm{Gamma}}
\newcommand{\dWei}{\mathrm{Wei}}
\newcommand{\dLogistic}{\mathrm{Logistic}}
\newcommand{\dRayleigh}{\mathrm{Rayleigh}}
\newcommand{\dPareto}{\mathrm{Pareto}}


% \renewcommand{\b}{\beta} % commented to avoid conflicts with base latex
% \renewcommand{\u}{u} % commented to avoid conflicts with base latex




% выделение питоновских и других пакетов
\newcommand{\sklearn}{sklearn}
\newcommand{\statsmodels}{statsmodels}


\usepackage{thmtools} % тонкая настройка окружения теорем


% окружения!
\declaretheorem[
style=definition, 
%thmbox=M,
title=Theorem, 
numberwithin=section, 
]{theorem}


\declaretheorem[
style=definition, 
title=Lemma, 
numberwithin=section, 
]{lemma}


\declaretheorem[
style=definition, 
title=Idea, 
% numberwithin=section, % ideas are rare!
]{idea} 



% опция sibling=theorem посередине даёт сквозную нумерацию с теоремами
\declaretheorem[
style=definition, 
title=Definition, 
sibling=theorem, 
]{definition}

\declaretheorem[
style=definition, 
title=Problem, 
% sibling=theorem, 
]{problem}

\declaretheorem[
style=definition, 
title=Exercise, 
% sibling=theorem, 
numberwithin=section,
]{exercise}


\declaretheorem[
style=definition,
title=Solution,
numbered=no
]{sol}


%\newenvironment{problem}{}{} % environment just prints the text
% \newenvironment{sol}{}{}
\newcommand{\trru}[1]{#1}
\newcommand{\tren}[1]{#1}



\begin{document}

% задачи можно дёргать:
% https://github.com/bdemeshev/metrics_pro/blob/master/metrics_body.tex
% и ещё (слегка модифицируя)
% https://github.com/bdemeshev/em_pset
% картиночки и доказательства на английском:
% https://github.com/olyagnilova/gauss-markov-pythagoras

\tableofcontents{}


\section{Day 1. Classic binomial and multinomial identities}

\subsection*{Planned entrance quiz :)}

\begin{enumerate}
\item How many ways are to order $6$ persons in a queue?
\item The box contains $50$ balls numbered from $1$ to $50$. We draw $6$ balls without replacement and write down the resulting sequence of numbers.
How many different sequences are possible?
\item I have $5$ yellow books and $7$ red books. I want to place them on my bookshelf and I would like to place yellow books to the left of red books. 
How many book orders are possible?
\end{enumerate}


\begin{definition}
    Factorial of $n$, denoted by $n!$, is the number of ways to put $n$ distinct objects in a queue. 
\end{definition}


\begin{definition}
    Binomial coefficient $n \choose k$ is the number of ways to select $k$ distinct objects from $n$ distinct objects.
\end{definition}

\begin{definition}
    For $n = a + b + c$ the multinomial coefficient $n \choose a, b, c$ is the number of ways to split $n$ distinct objects into different boxes, 
    where the first box contains $a$ objects, the second box contains $b$ objects, the third box contains $c$ objects. 

    One may create multinomial coefficient for any number of boxes :)
\end{definition}


\begin{exercise}
Is $n \choose k$ equal to $n \choose n-k$? 

Is $n \choose k$ equal to $n \choose k, n-k$?
\end{exercise}

\begin{exercise}
Write the product
    \[
    {100 \choose 20} \cdot {80 \choose 50} \cdot {30 \choose 10}
    \]
as one multinomial coefficient.
\end{exercise}

\begin{exercise}
Write 
\[
n \choose a, b, c, d
\] using binomial coefficients.

Write 
\[
n \choose a, b, c, d
\]
using factorials of $n$, $a$, $b$, $c$ and $d$.
\end{exercise}

\begin{idea}
Solution strategies for combinatorial identities:
\begin{enumerate}
    \item Double counting or story proof. 
    \item Visual identity (rare gem!).
    \item Evaluation of a polynomial at a special point. Differentiation, integration.
    \item Combine old identities into new. 
    \item Guess the answer and verify by induction. 
    \item Dark art!
\end{enumerate}
\end{idea}

\subsection*{Pascal's triangle and rule}
\begin{exercise}
Recall the Pascal arithmetic triangle. Draw two versions: triangular and rectangular. 
\end{exercise}

% Here the answer usually contains just one binomial coefficient or no at all:
\begin{exercise}
Pascal's rule. Simplify
\[
{n \choose k} + {n \choose k+1}.
\]
\[
{n \choose a-1, b, c} + {n \choose a, b-1, c} + {n \choose a, b, c-1}.
\]
\[
{n \choose a-1, b, c, d} + {n \choose a, b-1, c, d} + {n \choose a, b, c-1, d} + {n \choose a, b, c, d-1}.
\]
\end{exercise}

\begin{exercise}
Simplify 
\[
{n \choose 0} + {n \choose 1} + {n \choose 2} + \dots + {n \choose n}.
\]
Recall the Pascal's rule and simplify 
\[
{2n \choose 0} + {2n \choose 2} + {2n \choose 4} + \dots + {2n \choose 2n}.
\]
Simplify
\[
{2n-1 \choose 0} + {2n-1 \choose 2} + {2n-1 \choose 4} + \dots + {2n-1 \choose 2n-2}.
\]
\end{exercise}

\begin{exercise}
Simplify 
\[
\sum_{a + b + c = 2026} {2026 \choose a, b, c}.
\]    
\[
\sum_{a + b + c + d = 2027} {2027 \choose a, b, c, d}.
\]    
\[
\sum_{a + b + c = 100} {100 \choose a} \cdot {100-a \choose b}.
\]
\[
\sum_{a + b \leq n} {n \choose a} \cdot {n-a \choose b}.
\]
Here $a$, $b$ and $c$ are assumed non-negative integers.
\end{exercise}

\subsection*{Hockey-stick identity}

\begin{exercise}
Hockey-stick identity, Christmas stocking identity, boomerang identity.

Simplify
\[
\sum_{n = a}^b {n \choose a}.
\]
\[
1 + 2 + 3 + \dots + n.
\]
\[
1\cdot 2 + 2 \cdot 3 + 3 \cdot 4 + \dots + n \cdot (n+1).
\]
\[
1\cdot 2 \cdot 3 + 2 \cdot 3 \cdot 4 + 3 \cdot 4 \cdot 5 + \dots + n \cdot (n+1) \cdot (n+2).
\]
\end{exercise}

\begin{exercise}
Obtain closed form expressions for:     
\[
1^2 + 2^2 + \dots + n^2,
\]
\[
1^3 + 2^3 + \dots + n^3,
\]
\[
1^4 + 2^4 + \dots + n^4.
\]
\end{exercise}

\begin{exercise}
Hockey-sticks plus idea from linear algebra!

\begin{enumerate}[label=\alph*)]
\item Decompose $n^2 + 2n + 5$ using $n \choose 0$, $n \choose 1$, $n \choose 2$ as the basis. 

\item Find the sum 
\[
\sum_{k=2}^{2026} {k^2 + 2k + 5}.
\]

\item Using similar trick find the sums
\[
\sum_{k=1}^{n} {5k^2 + 4k}, \quad \sum_{k=1}^{n} {k^3 + 2k^2}.
\]
\end{enumerate}
\end{exercise}

\begin{exercise}
Let's start from some number $n \choose k$ inside the Pascal's triangle. 

In one step we can go either in north-east or north-west direction. 

Find the sum of all numbers we can reach.
\end{exercise}


\subsection*{Binomial or Taylor or MacLaurin expansion}
\begin{exercise}
Write the Taylor expansion (binomial expansion) at $x = 0$ for 
\[
(1 + x)^{n}, \quad (x + y)^n.
\]
\end{exercise}


\begin{exercise}
By plugging some special points to the $(x + y)^n$ or somewhere else find the sums
\[
{n \choose 0} + {n \choose 1} + {n\choose 2} + \dots + {n \choose n}.
\]
\[
\sum_{k=0}^n k {n \choose k}, \quad \sum_{k=0}^n (k-1) {n \choose k}, \quad \sum_{k=0}^n (-1)^k {n \choose k}, \quad \sum_{k=0}^n (-1)^{k-1} k {n \choose k}
\]

Hint: differentiation may help you.
\end{exercise}


\subsection*{How that can be?}

\begin{exercise}
Be brave, really brave! 

\begin{enumerate}[label=\alph*)]
\item Take a deep breath! 
\item Calculate 
\[
{-1 \choose 3}, \quad {-5 \choose 5},\quad  {1.5 \choose 3}, \quad  
{-1/2 \choose 4}, \quad  {1/3 \choose k}, \quad  {-1/2 \choose n}. 
\]
\end{enumerate}
\end{exercise}

\begin{exercise}
\begin{enumerate}[label=\alph*)]
\item Simplify 
\[
{-1/2 \choose 3} + {-1/2 \choose 4}.
\]

\item Draw the Pascal's triangle (the whole half-plane!) starting from $-1/2 \choose 0$. 

\item Draw the Pascal's triangle (the whole half-plane!) starting from $1/3 \choose 0$. 
\end{enumerate}
\end{exercise}


\begin{exercise}
Write the Taylor expansion at $x = 0$ for 
\[
\frac{1}{1 - x}, \quad  (1 + x)^{-5}, \quad  (1 + x)^{-1/2}, \quad  
(1 + 3x)^{1/3}, \quad  (1 - 5x)^{1/2}.
\]
\end{exercise}

\begin{exercise}
Simplify 
\[
{10 \choose 0}{-0.5 \choose 7} + {10 \choose 1}{-0.5 \choose 6} + {10 \choose 2}{-0.5 \choose 5} + \dots + {10 \choose 7}{-0.5 \choose 0}.
\]

Hint: for a honest proof consider the binomial expansion of $(1+x)^10$, $(1+x)^{-0.5}$, $(1+x)^{10-0.5}$.
\end{exercise}


\begin{exercise}
Homage to Newton. 

\begin{enumerate}[label=\alph*)]
\item Write the Taylor expansion at $x=0$ for $(1 - x^2)^{1/2}$.
\item Visually (or analytically if you wish) find the integral
\[
\int_{-1}^1 (1 - x^2)^{1/2}.
\]

\item Calculate the sums
\[
{1/2 \choose 0} - \frac{1}{3}{1/2 \choose 1} + \frac{1}{5} {1/2 \choose 2} + \dots
\]
\[
\sum_{k=2}^{\infty} \frac{1}{2k+1} \frac{(2k - 3)!!}{(2k)!!}
\]
\end{enumerate}
\end{exercise}



\subsection*{Home assignment-1:}

\begin{enumerate}
    \item Simplify
\[
\sum_{a + b + c = 100} {100 \choose a} \cdot {100-a \choose b}.
\]
    \item 1.9(2) in the old enumeration:

    Find explicit formula for the sum 
    \[
    1^3 + 2^3 + \dots + n^3.
    \]
    Hint: represent the general term $k^3$ as the sum of binomial coefficients and then use the hockey stick or Christmas stocking or boomerang!

    \item  {[*]} Using hockey stick decomposition find the sums
    \[
    \sum_{k=1}^{n} {5k^2 + 4k}.
    \]

    \item {[*]} Take a deep breath and calculate 
\[
{-1 \choose 3}, \quad {-5 \choose 5},\quad  {1.5 \choose 3}, \quad  
{-1/2 \choose 4}, \quad  {1/3 \choose k}, \quad  {-1/2 \choose n}. 
\]

    \item Write the Taylor expansion at $x = 0$ for 
\[
\frac{1}{1 - x}, \quad  (1 + x)^{-5}, \quad  (1 + x)^{-1/2}, \quad  
(1 + 3x)^{1/3}, \quad  (1 - 5x)^{1/2}.
\]

    \item Watch a Veritasium video «Newton changed the game»:

    \url{https://youtu.be/gMlf1ELvRzc?si=VASoVKQj3Pk-wstA}.
\end{enumerate}

You are more than welcome to solve other exercises that we skipped. Starred exercises will be on the quiz next day. 

\section{Day 2. Stars, bars and Gauss summation trick}

\subsection*{Day 2 welcome quiz!}

\begin{enumerate}
    \item Decompose the function $h(k) = 2k^2 + k$ using the binomial coefficient basis $k \choose 0$, $k \choose 1$ and $k \choose 2$.
    \item Let's play hockey! Find the sum 
    \[
    \sum_{k=2}^{1000} {3 {k \choose 0} - 2 {k \choose 1} + 3 {k \choose 2}}.
    \]
    Hint: do not forget to bring money back to the bank if you borrow!
    \item Find the value 
    \[
    {1/3 \choose 2} + {1/3 \choose 3} + {4/3 \choose 2}.
    \]
\end{enumerate}



\subsection*{Gauss summation trick}

\begin{exercise}
Simplify 
\[
1 + 2 + 3 + \dots + n;
\]
\[
1^2 + 2^2 + 3^2 + \dots + n^2;
\]
\[
{n \choose 1} + 2 {n \choose 2} + 3 {n \choose 3} + \dots + n{n \choose n};
\]
\[
{n \choose 0} + 2 {n \choose 1} + 3 {n \choose 2} + \dots + (n+1){n \choose n};
\]
\[
{n \choose 2} + 2 {n \choose 3} + 3 {n \choose 4} + \dots + (n-1){n \choose n};
\]
\[
{n \choose 0} + 3 {n \choose 1} + 5 {n \choose 2} + \dots + (2n+1){n \choose n};
\]
\[
{n \choose 0} - 2 {n \choose 1} + 3 {n \choose 2} - 4 {n \choose 3} \dots + (-1)^n (n+1){n \choose n};
\]
\[
3{n \choose 1} + 7 {n \choose 2} + 11 {n \choose 3} + \dots + (4n-1){n \choose n};
\]
\[
{n \choose 1} - 2 {n \choose 2} + 3 {n \choose 3} - 4 {n \choose 4} + \dots + (-1)^n n{n \choose n};
\]
\end{exercise}


\subsection*{Stars and bars}

\begin{exercise}
There are $20$ chairs in a row and $7$ students. Each student needs one chair to sit down. 
We do care only about which chairs are occupied.

\begin{enumerate}[label=\alph*)]
    \item In how many ways chairs can be occupied? 
    \item In how many ways chairs can be occupied if there should be at least one empty chair between occupied chairs?
    \item In how many ways chairs can be occupied if there should be at least two empty chairs between occupied chairs?
\end{enumerate}
\end{exercise}

\begin{exercise}
I have four bags of different colors and $20$ identical marble stones. 

\begin{enumerate}[label=\alph*)]
    \item In how many different ways can I put marbles into bags?
    \item In how many different ways can I put marbles into bags if there should be no empty bags?
\end{enumerate}
\end{exercise}

\begin{theorem}[Stars and bars theorem]
The number of ways to distribute $n$ identical objects into $k$ distinguishable bags is equal to
\[
{n + k - 1 \choose n} = {n + k - 1 \choose k - 1}.
\]
Hint: $n$ is the number of stars and $(k - 1)$ is the number of bars. 
\end{theorem}

\begin{exercise}
There are $5$ types of post-cards sold at the post-office. 
\begin{enumerate}[label=\alph*)]
\item In how many different ways can I buy $12$ post-cards?
\item In how many different ways can I buy $12$ post-cards if I would like to buy at least one post-card of every type?
\end{enumerate}
\end{exercise}

\begin{exercise}
Consider the equation $x_1 + x_2 + x_3 + x_4 = 2026$.

\begin{enumerate}[label=\alph*)]
\item How many solutions does it have if $x_1$, $x_2$, $x_3$ and $x_4$ are non-negative integers?
\item How many solutions does it have if $x_1$, $x_2$, $x_3$ and $x_4$ are positive integers?
\item How many solutions does it have if all $x_i$ are integers and $x_i \geq 2$ for all $i$?
\item How many solutions does it have if all $x_i$ are integers and $x_i \geq i$ for all $i$?
\end{enumerate}
\end{exercise}


\begin{exercise}
There are $10$ standard indistinguishable dices. 

\begin{enumerate}[label=\alph*)]
    \item How many different combinations of scores are possible?
    \item How many different combinations of scores are possible if at least one dice shows $6$?
    \item How many different combinations of scores are possible if no dice shows $5$ nor $6$?
\end{enumerate}
\end{exercise}

\begin{exercise}
Stars, bars, circles and probability!

Consider a well shuffled deck of $52$ cards. I open the cards one by one. 

\begin{enumerate}[label=\alph*)]
    \item How many cards on average I open to see the first Queen?
    \item How many cards on average I open to see the third Ace?
    \item What is expected number of cards between the third and the fourth King?
    \item What is the probability that the third Ace will be the $20$-th card opened?
    \item Write combinatorial identities hidden in a), b) and c).
\end{enumerate}
    
\end{exercise}

\section*{Visualize it}

\begin{exercise}
\begin{enumerate}[label=\alph*)]
\item Express as a binomial coefficient and visualize the sum:
\[
1 + 2 + 3 + 4 + \dots + n.
\]
\item Visualize the Fuji-sum and simplify it!
\[
1 + 2 + 3  + \dots + (n-1) + n + (n-1) + \dots + 3 + 2 + 1.
\]
% pull-ups ladder?
\item Recall the multiplication table for numbers from $1$ to $10$.
Why the sum of all numbers is just 
\[
1^3 + 2^3 + 3^3 + \dots + 10^3?
\]
Simplify the sum!
\end{enumerate}

\end{exercise}

\subsection*{Home assignment: day 2 $\to$ day 5}

Deadline: 2026-07-03, 16:59. 

\begin{enumerate}
    \item There are $10$ standard indistinguishable dices. 

\begin{enumerate}[label=\alph*)]
    \item How many different combinations of scores are possible?
    \item How many different combinations of scores are possible if at least one dice shows $6$?
    \item How many different combinations of scores are possible if no dice shows $5$ nor $6$?
\end{enumerate}

\item There are $5$ types of post-cards sold at the post-office. 
\begin{enumerate}[label=\alph*)]
\item In how many different ways can I buy $12$ post-cards?
\item In how many different ways can I buy $12$ post-cards if I would like to buy at least one post-card of every type?
\end{enumerate}

\item Using Gauss summation trick or otherwise find the sum:
\[
{n \choose 0} - 2 {n \choose 1} + 3 {n \choose 2} - 4 {n \choose 3} \dots + (-1)^n (n+1){n \choose n};
\]

\item Recall the mother-identity
\[
(1 + x)^n = 1 + {n \choose 1} x^1 + {n \choose 2} x^2 + \dots + {n \choose n} x^n.
\]

\begin{enumerate}
    \item Find the first derivative of the mother-identity. 
    \item Find the value of the sum
\[
{n \choose 1} + 2 \cdot 7 \cdot {n \choose 2} + 3 \cdot 7^2 \cdot {n \choose 2} + \dots + n \cdot 7^{n-1} {n \choose n}.
\]
\item Find the second derivative of the mother-identity.
\item Find the value of the sum
\[
2 \cdot 1 {n \choose 2} + 3 \cdot 2 \cdot 9 \cdot {n \choose 3} + 4 \cdot 3 \cdot 9^2 \cdot {n \choose 4} + 5 \cdot 4 \cdot 9^3 \cdot {n \choose 5} + \dots + n(n-1) 9^{n-2} {n \choose n}.
\]

\end{enumerate}

\end{enumerate}



\subsection*{Demo for the quiz on day 3}

\begin{enumerate}

    \item Using Gauss summation trick or otherwise find the sum 
    \[
    3 {n \choose 1} + 5 {n \choose 2} + 7 {n \choose 3} + \dots + (2n + 1) {n \choose n}.
    \]
    \item Consider the equation $a + b + c = 100$ with integers $a \geq 3$, $b \geq 2$, $c\geq 5$.

    How many solutions are there?
    
\end{enumerate}

\section{Day 3. Recurrence relations}

\begin{comment}
\subsection*{Quiz day 3}

\begin{enumerate}

    \item Using Gauss summation trick or otherwise find the sum 
    \[
    -5{n\choose 0} -3 {n \choose 1} -1 \cdot {n \choose 2} + 1 \cdot {n \choose 3} + \dots + (2n - 5) {n \choose n}.
    \]
    \item Consider the equation $a + b + c = 10$ with integers $a \geq 2$, $b \geq 1$, $c\geq 3$.

    \begin{enumerate}
        \item How many solutions are there?
        \item Which solution would you encode as $\star \mid \, \mid \star\star\star$?
    \end{enumerate}
\end{enumerate}
\end{comment}


\begin{exercise}
\begin{enumerate}[label=\alph*)]
    Recurrence equations starter pack :)
    
    \item Solve the recurence equation:
    \[
    a_n = a_{n-1} + 3, \quad a_1 = 20.
    \]

    \item Solve the recurence equation:
    \[
    b_n = 3 \cdot b_{n-1}, \quad b_1 = 20.
    \]
    \item Solve the recurence equation:
    \[
    c_n = (n+3) \cdot c_{n-1}, \quad c_1 = 20.
    \]
    \item Solve the recurence equation:
    \[
    d_n = n+3  + d_{n-1}, \quad d_1 = 20.
    \]
    \item Solve the recurrence equation:
    \[
    e_n = n^2 - 3n + 2  + e_{n-1}, \quad e_1 = 20.
    \]
    \item Solve the recurrence equation:
    \[
    f_n = (n^2 - 3n + 2) f_{n-1}, \quad f_1 = 20.
    \]
    \item Solve the recurrence equation:
    \[
    g_n = \frac{n+2}{n+1} g_{n-1}, \quad g_1 = 20.
    \]
    \item Solve the recurrence equation:
    \[
    h_n = \frac{1}{n^2 + n} + h_{n-1}, \quad h_1 = 20.
    \]
\end{enumerate}
\end{exercise}

\begin{exercise}
    Gauss trick with substraction!
    \begin{enumerate}[label=\alph*)]
        \item Find the sum
        \[
        S = 2 + 2 \cdot (1/3) + 2 \cdot (1/3)^2 + 2 \cdot (1/3)^3 + \dots
        \]
        \item Solve the recurrence equation 
        \[
        S_n = S_{n-1} + 3 \cdot (1/5)^n, \quad S_0 = 0.
        \]
        \item Solve the recurrence equation 
        \[
        S_n = S_{n-1} + n \cdot (1/5)^n, \quad S_0 = 0.
        \]
    \end{enumerate}

\end{exercise}

\begin{exercise}
You draw $n$ lines on a plane, they split the plane into some number of regions. 
\begin{enumerate}[label=\alph*)]
    \item What is the minimal number of regions?
    \item What is the maximal number of regions?
\end{enumerate}
\end{exercise}


\begin{idea}
    Consider the equation 
    \[
    x_t + \alpha_1 x_{t-1} + \alpha_2 x_{t-2} = \textrm{poly(t)}.
    \]
    The solution can be always represented as a sum or product of some geometric series and polynomials.
\end{idea}

% idea: first we start by guessing polynomial solution and then we simplify to homogeneous!

\begin{exercise}
Unique root case. 
\begin{enumerate}[label=\alph*)]
    \item Solve the recurence equation:
    \[
    a_n - 5 a_{n-1} + 6 a_{n-2} = 0, \quad a_0 = 3, a_1 = 5.
    \]
    \item Solve the recurence equation:
    \[
    a_n - 7 a_{n-1} + 12 a_{n-2} = 6, \quad a_0 = 1, a_1 = 12.
    \]
    \item Solve the recurence equation:
    \[
    a_n - 8 a_{n-1} + 15 a_{n-2} = 8n - 22 , \quad a_0 = 5, a_1 = 22.
    \]
\end{enumerate}
\end{exercise}

\begin{exercise}
    % Consider the Fibonacci numbers $F_n$.
    We start with one pair of young rabbits $F_1 = 1$. 
    It takes one period of time for a pair of rabbits to mature. 
    Every mature pair of rabbits in one period of time will give birth to a new young pair of rabbits. 

    \begin{enumerate}[label=\alph*)]
        \item Find $F_2$, $F_3$, $F_4$, $F_5$.
        \item Deduce the recurrence relation for $F_n$.
        \item Find the general formula for $F_n$.
        \item Which term in the general formula can be ignored for the big $n$?
        \item Guess and prove formulas for 
        \[
        \sum_{k=0}^n F_k, \quad \sum_{k=0}^n F_{2k}, \quad \sum_{k = 1}^n F_{2k-1}, \quad F_{n+1} F_{n-1} - F_n^2.
        \]
        \item Which of the following formulas are valid?
        \[
        F_{m + n} = F_m F_{n+1} + F_{m-1} F_n
        \]
        \[
        F_n^2 + F_{n+1}^2 = F_{2n+1}
        \]
        \[
        F_{n+1}^2 - F_{n-1}^2 = F_{2n}
        \]
        \[
        \sum_{k=0}^n F_k^2 = F_n F_{n+1}.
        \]
    \end{enumerate}

\end{exercise}


\begin{exercise}
    The mighty lag operator!

    Let $L$ be the lag operator and $\Delta = 1 - L$. Consider the sequences $x_n = 5^n$, $y_n = n\cdot 5^n$ and $z_n = 3^n$.
    \begin{enumerate}[label=\alph*)]
        \item Find $(1 - 3L)x_n$, $\Delta x_n$, $(1 - 5L)x_n$.
        \item Find $\Delta (1 - 3L) x_n$, $\Delta (1- 5L)x_n$, $(1 - 5L)^2 x_n$.
        \item Find $(1 - 3L)y_n$, $\Delta y_n$, $(1 - 5L)y_n$.
        \item Find $\Delta (1 - 3L) y_n$, $\Delta (1- 5L)y_n$, $(1 - 5L)^2 y_n$.
        \item Find $(1 - 3 L)(1 - 5L) x_n$ and $(1 - 3L)(1 - 5L) z_n$.
        \item Rewrite the equation $a_t - 7a_{t-1} + 12 a_{t-2} = 0$ using lag operator. 
    \end{enumerate}

Warning! Remember that $(1 - 3L)$ or $(1 - 5L)$ is not a number! Lag operator is just a convinient notation to shift index!
\end{exercise}


\begin{exercise}
Double root case. 
\begin{enumerate}[label=\alph*)]
    \item Solve the recurence equation:
    \[
    a_n - 8 a_{n-1} + 16 a_{n-2} = 0, \quad a_0 = 3, a_1 = 0.
    \]
    \item Solve the recurence equation:
    \[
    a_n - 4 a_{n-1} + 13 a_{n-2} = 20, \quad a_0 = 6, a_1 = 4.
    \]
    \item Solve the recurence equation:
    \[
    q_n - 6 q_{n-1} + 9q_{n-2} = 8n - 20, \quad q_0 = 2, a_1 = 12.
    \]
\end{enumerate}

\end{exercise}


\subsection*{Home assignment: day 3 $\to$ day 6}

Deadline: 2026-07-06, 14:59. 

\begin{enumerate}
    \item Solve the recurrence equation: 
    \[
    a_n = (n + 1) + a_{n-1}, \quad a_1 = 100.
    \]
    Hint: recall that $1 + 2 + 3 + 4 + \dots + n = {n+1 \choose 2} = \frac{n(n+1)}{2}$.
    \item Solve the recurence equation:
    \[
    a_n + 5 a_{n-1} + 6 a_{n-2} = 0, \quad a_0 = 3, a_1 = 5.
    \]
    \item Solve the recurence equation:
    \[
    c_n = 2 \cdot c_{n-1} + 3, \quad c_1 = 20.
    \]
    \item Solve the recurence equation:
    \[
    a_n - 7 a_{n-1} + 12 a_{n-2} = 6, \quad a_0 = 1, a_1 = 12.
    \]
     \item Watch a Veritasium video «This equation will change how you see the world»:

    \url{https://youtu.be/ovJcsL7vyrk?si=kh8wjM2e6dgLSXzj}.
\end{enumerate}


\subsection*{Demo for the quiz on day 4}

\begin{enumerate}
    \item Solve the recurrence equation:
    \[
    b_n = \frac{n + 5}{n + 4} b_{n-1}, \quad b_1 = 100.
    \]
    \item Here $L$ is the lag operator and $x_n = 6^n$. 
    Find the values
    \[
    (1 - 5L) x_n, \quad (1 - 6L)x_n, \quad (1-6L)(1 - 5L)x_n.
    \]
\end{enumerate}



\section{Day 4. More recurence equations}

\begin{comment}
\subsection*{Quiz day 4}

\begin{enumerate}
    \item Solve the recurrence equation:
    \[
    b_n = \frac{n + 2}{n + 3} b_{n-1}, \quad b_1 = 10.
    \]
    \item Here $L$ is the lag operator and $x_n = 2^n$. 
    Find the values
    \[
    (1 - 3L) x_n, \quad (1 - 2L)x_n, \quad (1-2L)(1 - 3L)x_n.
    \]
\end{enumerate}


\end{comment}


\begin{exercise}
Double root case. 
\begin{enumerate}[label=\alph*)]
    \item Solve the recurence equation:
    \[
    a_n - 8 a_{n-1} + 16 a_{n-2} = 0, \quad a_0 = 3, a_1 = 0.
    \]
    \item Solve the recurence equation:
    \[
    a_n - 4 a_{n-1} + 4 a_{n-2} = 20, \quad a_0 = 20, a_1 = 21.
    \]
    \item Solve the recurence equation:
    \[
    q_n - 6 q_{n-1} + 9q_{n-2} = 8n - 20, \quad q_0 = 2, a_1 = 12.
    \]
\end{enumerate}

\end{exercise}



\begin{exercise}
Unique complex root case. 
\begin{enumerate}[label=\alph*)]
    \item Solve the recurence equation:
    \[
    a_n - 4 a_{n-1} + 13 a_{n-2} = 0, \quad a_0 = 4, a_1 = 2.
    \]
    \item Solve the recurence equation:
    \[
    a_n - 4 a_{n-1} + 13 a_{n-2} = 20, \quad a_0 = 6, a_1 = 4.
    \]
    \item Solve the recurence equation:
    \[
    a_n + 9 a_{n-2} = 10n - 8, \quad a_0 = 7, a_1 = 8.
    \]
\end{enumerate}

\end{exercise}


\begin{exercise}
There are $n$ passengers boarding an intercity bus and standing in queue. 
There are $n$ places and each passenger has its own place on the ticket. 
The problem is that the first passenger in the queue is very nervous, he rushes in the bus and takes one place at random. 
Each next passenger takes his seat according to the ticket if the seat is free. 
And if his own seat is occupied each passenger takes a random free seat. 

Let $p_n$ be the probability that the last passenger in the queue will get his seat. 
\begin{enumerate}[label=\alph*)]
    \item What is the initial conditions $p_{2}$?
    \item Build the recurrence equation for $p_{n}$.
    \item Guess the solution!
\end{enumerate}
\end{exercise}

\begin{exercise}
I have $100$ safes. Each safe can be closed without a key, but requires a specific unique key to be opened. 
I randomly distribute $100$ keys into $100$ safes, one key per safe. 
I leave $5$ safes open and I close other safes. So I have access to $5$ keys and I can use them to open other safes and inside these safes there are more keys\dots{ }

Let's denote by $p_{kn}$ the probability that I can open all $n$ safes, given that in the beginning there are $k$ safes open and $(n-k)$ safes closed.
\begin{enumerate}[label=\alph*)]
    \item What are the initial conditions $p_{0n}$ and $p_{nn}$?
    \item Build the recurrence equation for $p_{kn}$.
    \item Guess the solution!
\end{enumerate}
Let's denote by $e_{kn}$ the expected number of open safes, given that in the beginning there are $k$ safes open and $(n-k)$ safes closed.
\begin{enumerate}[resume]
    \item What are the initial conditions $e_{0n}$ and $e_{nn}$?
    \item Build the recurrence equation for $e_{kn}$.
    \item Guess the solution!
\end{enumerate}


\end{exercise}


\begin{exercise}
I enter a casino in Monte Carlo with $50$ tokens. I play the roulette.
I always bet one token on black. I will leave the casino when my fortune will be $70$ tokens or $0$ tokens. 

In a typical Monte Carlo roulette there are $18$ black pockets, $18$ red pockets and zero. 
Let $p_n$ be the probability that I will leave the casino as winner if I have $n$ tokens now. 
\begin{enumerate}[label=\alph*)]
    \item What are the initial conditions $p_0$ and $p_{70}$?
    \item Build the recurrence equation for $p_n$. 
    \item Solve the recurrence equation and find $p_{50}$.
    \item How the answers would change for Las Vegas roulette with zero and double zero pockets?
    \item How the answers would change for an abstract fair casino with no zero nor double zero?
\end{enumerate}
Let $e_n$ be the expected number of spins before I leave the casino (as winner or as loser).
\begin{enumerate}[resume]
    \item What are the initial conditions $e_0$ and $e_{70}$?
    \item Build the recurrence equation for $e_n$. 
    \item Solve the recurrence equation and find $e_{50}$.
    \item How the answers would change for Las Vegas roulette with zero and double zero pockets?
    \item How the answers would change for an abstract fair casino with no zero nor double zero?
\end{enumerate}

    


\end{exercise}


\begin{exercise}
I have a broken light switch button in my room. 
The switch button should turn the light on if the light is off and vice versa.
On the first press the button works with probability $1/2$, on the second press — with probability $1/3$, on the third one — with probability $1/4$ and so on.
Initially my room is dark.

What is the probability that my room will be dark after I press the button $100$ times?
\end{exercise}



\begin{exercise}
You play a the following lottery with the casino. There is a hat on the table. 
A fair dice is thrown until the game ends. 
\begin{itemize}
    \item If the dice shows one then you get $+1$ dollar immediately and the game continues. 
    \item If the dice shows two then $+2$ dollars are added to the hat and the game continues. 
    \item If the dice shows three then $+3$ dollars are added to the hat and the game continues.
    \item If the dice shows four then the sum in the hat goes back to casino and the game continues.
    \item If the dice shows five then the sum in the hat goes back to the casino and the game ends.
    \item If the dice shows six then the game ends and you receive the content of the hat. 
\end{itemize}

Let $e_n$ be your total expected payoff if there are $n$ dollars in the hat now. 

\begin{enumerate}
    \item Build the recurrence equation for $e_n$.
    \item Find the solution.
    \item What is the price of one dollar in the hat in real dollars? What is the fair price to participate in this game?
\end{enumerate}
\end{exercise}


\begin{exercise}
    Now the systems!

    \begin{enumerate}[label=\alph*)]
    \item Solve the recurrence equation:
    \[
    \begin{cases}
    a_n = 3a_{n-1} + 2b_{n-1} - 4, \\
    b_n = 2a_{n-1} + 3b_{n-1} - 4, \\
    a_0 = 1, \quad b_0 = 0.
    \end{cases}
    \]
    \item Solve the recurrence equation:
    \[
    \begin{cases}
    a_n = 3a_{n-1} - 2b_{n-1} - 4, \\
    b_n = - 2a_{n-1} + 3b_{n-1} + 4, \\
    a_0 = 1, \quad b_0 = 0.
    \end{cases}
    \]
    \end{enumerate}

\end{exercise}


\section{Day 5. Binomial coefficients and recurrence relations recap}




\section{Yet to discover}


\subsection*{Inclusions and exclusions}


\subsection*{Proofs to DIE for}



\subsection*{Wandermonde's identity}

\begin{exercise}
Simplify 
\[
{a \choose 0}{b \choose m} + {a \choose 1}{b \choose m-1} + {a \choose 2}{b \choose m-2} + \dots + {a \choose m}{b \choose 0}.
\]
\[
{a \choose 0}{m \choose 0} + {a \choose 1}{m \choose 1} + {a \choose 2}{m \choose 3} + \dots + {a \choose m}{m \choose m}.
\]
\[
{a \choose 0}^2 + {a \choose 1}^2  + {a \choose 2}^2 + \dots + {a \choose a}^2.
\]
\end{exercise}

\begin{exercise}
There are $n = 500$ fishes in the pond, $t = 30$ of them are tagged. You capture $c = 5$ fishes at random without replacement. 
Let the random variable $Y$ be the number of captured tagged fishes. 
Find the probabilities $\P(Y = 0)$, $\P(Y = 1)$, $\P(Y = 2)$, $\P(Y = y)$.
\end{exercise}

\begin{exercise}
There are $n = 24$ candies in the jar, $14$ with orange flavour and $10$ with ananas flavour.
Your take $6$ candies at random. Let $Y$ be the number of candies with orange flavour you picked. 
Find the probabilities $\P(Y = 0)$, $\P(Y = 1)$, $\P(Y = 2)$, $\P(Y = y)$.
\end{exercise}

\begin{exercise}
A bridge hand (random $13$ cards out of $52$-card standard deck) is dealt. 

\begin{enumerate}[label=\alph*)]
\item What is the probability that the hand contains no spades? 
\item What is the probability that the hand contains exactly $5$ hearts? 
\item What is the probability that the hand contains at most two aces? 
\end{enumerate}
\end{exercise}


\begin{exercise}
Wandermonde-like identities in a funny notation.
Here I intentionally omit the running index $i$, I write the first summation term and specify with $\uparrow$ and $\downarrow$ which terms goes up or down in the sum. 
\begin{enumerate}[label=\alph*)]
\item Simplify and rewrite in a standard notation
% n persons: k to chill, (n-k) to work on two types of work: A, B
\[
\sum {n \choose 0^{\uparrow}} {n^{\downarrow} \choose k}.
\]
\item Simplify and rewrite in a standard notation
% a + n + 1 chairs in total: a hard working students, b lazy students and 1 TA. 
% we put them in a specified order hard working students, TA, lazy students. 
% {a+n+1 \choose a + b + 1} ways in total
% sum is split by the number of chairs before and after TA
\[
\sum {a^{\uparrow} \choose a} {n^{\downarrow} \choose k}.
\]
\item Simplify and rewrite in a standard notation
% (n+b) persons: b boys, n girls, we need to choose k persons.
\[
\sum {b \choose 0^{\uparrow}} {n \choose k^{\downarrow}}.
\]
\end{enumerate}
\end{exercise}



\begin{exercise}
Simplify 
\[
\sum_{a + b + c = k}{n_1 \choose a}{n_2 \choose b}{n_3 \choose c}. % {n_1 + n_2 + n_3 \choose k}
\]
\end{exercise}



\nocite{graham1994concrete}
\nocite{harris2010combinatorics}
\nocite{guichard2025introcgt}    


\newpage
\addcontentsline{toc}{section}{Sources of wisdom}
\printbibliography[heading=none]


\end{document}
