% arara: xelatex
\documentclass[12pt]{article}

% \usepackage{physics}

\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
    pdftitle={Overleaf Example},
    pdfpagemode=FullScreen,
    }

\usepackage{tikzducks}

\usepackage{tikz} % картинки в tikz
\usetikzlibrary{shapes, arrows, positioning}
\usepackage{microtype} % свешивание пунктуации

\usepackage{array} % для столбцов фиксированной ширины

\usepackage{indentfirst} % отступ в первом параграфе

\usepackage{sectsty} % для центрирования названий частей
\allsectionsfont{\centering}

\usepackage{amsmath, amsfonts, amssymb} % куча стандартных математических плюшек

\usepackage{comment}

\usepackage[top=2cm, left=1.2cm, right=1.2cm, bottom=2cm]{geometry} % размер текста на странице

\usepackage{lastpage} % чтобы узнать номер последней страницы

\usepackage{enumitem} % дополнительные плюшки для списков
%  например \begin{enumerate}[resume] позволяет продолжить нумерацию в новом списке
\usepackage{caption}

\usepackage{url} % to use \url{link to web}

\usepackage{xcolor}   % для цвета текста
\usepackage{ulem}     % для зачёркивания (\sout)

\usepackage[symbol]{footmisc}

\newcommand{\smallduck}{\begin{tikzpicture}[scale=0.3]
    \duck[
        cape=black,
        hat=black,
        mask=black
    ]
    \end{tikzpicture}}

\usepackage{fancyhdr} % весёлые колонтитулы
\pagestyle{fancy}
\lhead{Harbour.space}
\chead{}
\rhead{}
\lfoot{}
\cfoot{}
\rfoot{}

\renewcommand{\headrulewidth}{0.4pt}
\renewcommand{\footrulewidth}{0.4pt}

\usepackage{tcolorbox} % рамочки!

\usepackage{todonotes} % для вставки в документ заметок о том, что осталось сделать
% \todo{Здесь надо коэффициенты исправить}
% \missingfigure{Здесь будет Последний день Помпеи}
% \listoftodos - печатает все поставленные \todo'шки


% более красивые таблицы
\usepackage{booktabs}
% заповеди из докупентации:
% 1. Не используйте вертикальные линни
% 2. Не используйте двойные линии
% 3. Единицы измерения - в шапку таблицы
% 4. Не сокращайте .1 вместо 0.1
% 5. Повторяющееся значение повторяйте, а не говорите "то же"


\setcounter{MaxMatrixCols}{20}
% by crazy default pmatrix supports only 10 cols :)


\usepackage{fontspec}
\usepackage{libertine}
\usepackage{polyglossia}

\setmainlanguage{russian}
\setotherlanguages{english}


\usetikzlibrary{automata, positioning, arrows.meta, calc, shapes.multipart}


% download "Linux Libertine" fonts:
% http://www.linuxlibertine.org/index.php?id=91&L=1
% \setmainfont{Linux Libertine O} % or Helvetica, Arial, Cambria
% why do we need \newfontfamily:
% http://tex.stackexchange.com/questions/91507/
% \newfontfamily{\cyrillicfonttt}{Linux Libertine O}

\AddEnumerateCounter{\asbuk}{\russian@alph}{щ} % для списков с русскими буквами
% \setlist[enumerate, 2]{label=\asbuk*),ref=\asbuk*}

%% эконометрические сокращения
\DeclareMathOperator{\Cov}{\mathbb{C}ov}
\DeclareMathOperator{\Corr}{\mathbb{C}orr}
\DeclareMathOperator{\Var}{\mathbb{V}ar}
\DeclareMathOperator{\col}{col}
\DeclareMathOperator{\row}{row}

\let\P\relax
\DeclareMathOperator{\P}{\mathbb{P}}

\DeclareMathOperator{\E}{\mathbb{E}}
% \DeclareMathOperator{\tr}{trace}
\DeclareMathOperator{\card}{card}

\DeclareMathOperator{\Convex}{Convex}
\DeclareMathOperator{\plim}{plim}

\newcommand{\cN}{\mathcal{N}}
\newcommand{\cF}{\mathcal{F}}

\newcommand{\RR}{\mathbb{R}}
\newcommand{\NN}{\mathbb{N}}
\newcommand{\hb}{\hat{\beta}}
\newcommand{\dPois}{\mathrm{Pois}}
\newcommand{\dUnif}{\mathrm{Unif}}

\usepackage{mathtools}


\DeclarePairedDelimiter{\norm}{\lVert}{\rVert}
\DeclarePairedDelimiter{\abs}{\lvert}{\rvert}
\DeclarePairedDelimiter{\scalp}{\langle}{\rangle}
\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}




\begin{document}

%Madame de Pompadour%\footnote{official chief mistress of King Louis XV} 
%was born exactly 204 years earlier, on 29 December 1721. 
%Her original name was Jeanne Antoinette Poisson.





\subsection*{Final demo :)}

This is a demo version of the final exam. 
Final will consist of 6 problems: generating functions in probability, Polya-Redfield counting, planarity and Euler formula, graph colouring, linearity of expected value and one old topic.
% midterm: two problems on binomial identities, two problems on recurrent equations, two problems on generating functions. 
Some problems will be similar to demo \begin{tikzpicture}[scale=0.3]
     \shuffleducks[13]
    \randuck
\end{tikzpicture}

You may use one A4 cheat sheet in any form handwritten, printed, aquarelle painting\dots{ } 

\begin{enumerate}
    \item {[old topic]} Write the Taylor expansion at $x_0 = 0$ of the function 
    \[
    f(x) = (1 - 5x^2)^{1/2}.
    \]
    \item {[generating functions in probability]} 
    I throw a coin until the sequence $HTT$ appears. The coin is not fair and shows $H$ with probability $0.4$.
\begin{enumerate}[label=\alph*)]
    \item Draw the directed graph for this game.
    \item Setup a system of equations for the sequences with given endpoint and starting point at the start of the game. 
    \item Explain how would you calculate expected number of coin tosses from the start till the end of the game.
    
    Note: you don't need to do actual calculations nor to solve the system, just describe how can you extract the expected value using derivatives. 
\end{enumerate}

    \item {[Polya-Redfield counting]} Consider a $2\times 2$ mosaic of gems, where each gem may be of one of $k$ possible types.
    We consider mosaics identical if the can be transformed one into another by rotation or flip. 
\begin{enumerate}[label=\alph*)]
    \item List all elements of the group of actions that act on mosaics.
    \item Write each action as a combination of cycles and provide its cyclic monomial. 
    \item Find the cyclic index and the number of different possible colourings.
\end{enumerate}
    \item {[planarity and Euler formula]} The convex polyhedron consists only of  triangular and  hexagonal faces  with  three faces meeting at each vertex.
    
    How many triangular faces are there?

    \item {[graph colouring]} Consider the graph with two components. The first component is a $K_4$ graph, the second one — $C_4$ graph. 
    Components are non connected between them. You have $k$ colours available.

    In how many ways can you color the vertices of this graph if adjacent colours can't be the same?
    
    
    \item {[linearity of expected value]} There are $20$ types of desserts in a café. In the queue there are $10$ clients.
Each client prefers one of the desserts at random with equal probabilities.

\begin{enumerate}[label=\alph*)]
  \item What is the average number of dessert types bought?
  \item What is the average number of dessert types bought if each client will not buy any dessert with probability $0.3$?
\end{enumerate}

\end{enumerate}


\end{document}

