% arara: xelatex
\documentclass[12pt]{article}

% \usepackage{physics}

\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
    pdftitle={Overleaf Example},
    pdfpagemode=FullScreen,
    }

\usepackage{tikzducks}

\usepackage{tikz} % картинки в tikz
\usetikzlibrary{shapes, arrows, positioning}
\usepackage{microtype} % свешивание пунктуации

\usepackage{array} % для столбцов фиксированной ширины

\usepackage{indentfirst} % отступ в первом параграфе

\usepackage{sectsty} % для центрирования названий частей
\allsectionsfont{\centering}

\usepackage{amsmath, amsfonts, amssymb} % куча стандартных математических плюшек

\usepackage{comment}

\usepackage[top=2cm, left=1.2cm, right=1.2cm, bottom=2cm]{geometry} % размер текста на странице

\usepackage{lastpage} % чтобы узнать номер последней страницы

\usepackage{enumitem} % дополнительные плюшки для списков
%  например \begin{enumerate}[resume] позволяет продолжить нумерацию в новом списке
\usepackage{caption}

\usepackage{url} % to use \url{link to web}

\usepackage{xcolor}   % для цвета текста
\usepackage{ulem}     % для зачёркивания (\sout)

\usepackage[symbol]{footmisc}

\newcommand{\smallduck}{\begin{tikzpicture}[scale=0.3]
    \duck[
        cape=black,
        hat=black,
        mask=black
    ]
    \end{tikzpicture}}

\usepackage{fancyhdr} % весёлые колонтитулы
\pagestyle{fancy}
\lhead{Harbour.space}
\chead{Final exam}
\rhead{\begin{tikzpicture}[scale=0.3]
     \shuffleducks[13]
    \randuck
\end{tikzpicture}
}
\lfoot{}
\cfoot{}
\rfoot{}

\renewcommand{\headrulewidth}{0.4pt}
\renewcommand{\footrulewidth}{0.4pt}

\usepackage{tcolorbox} % рамочки!

\usepackage{todonotes} % для вставки в документ заметок о том, что осталось сделать
% \todo{Здесь надо коэффициенты исправить}
% \missingfigure{Здесь будет Последний день Помпеи}
% \listoftodos - печатает все поставленные \todo'шки


% более красивые таблицы
\usepackage{booktabs}
% заповеди из докупентации:
% 1. Не используйте вертикальные линни
% 2. Не используйте двойные линии
% 3. Единицы измерения - в шапку таблицы
% 4. Не сокращайте .1 вместо 0.1
% 5. Повторяющееся значение повторяйте, а не говорите "то же"


\setcounter{MaxMatrixCols}{20}
% by crazy default pmatrix supports only 10 cols :)


\usepackage{fontspec}
\usepackage{libertine}
\usepackage{polyglossia}

\setmainlanguage{russian}
\setotherlanguages{english}


\usetikzlibrary{automata, positioning, arrows.meta, calc, shapes.multipart}


% download "Linux Libertine" fonts:
% http://www.linuxlibertine.org/index.php?id=91&L=1
% \setmainfont{Linux Libertine O} % or Helvetica, Arial, Cambria
% why do we need \newfontfamily:
% http://tex.stackexchange.com/questions/91507/
% \newfontfamily{\cyrillicfonttt}{Linux Libertine O}

\AddEnumerateCounter{\asbuk}{\russian@alph}{щ} % для списков с русскими буквами
% \setlist[enumerate, 2]{label=\asbuk*),ref=\asbuk*}

%% эконометрические сокращения
\DeclareMathOperator{\Cov}{\mathbb{C}ov}
\DeclareMathOperator{\Corr}{\mathbb{C}orr}
\DeclareMathOperator{\Var}{\mathbb{V}ar}
\DeclareMathOperator{\col}{col}
\DeclareMathOperator{\row}{row}

\let\P\relax
\DeclareMathOperator{\P}{\mathbb{P}}

\DeclareMathOperator{\E}{\mathbb{E}}
% \DeclareMathOperator{\tr}{trace}
\DeclareMathOperator{\card}{card}

\DeclareMathOperator{\Convex}{Convex}
\DeclareMathOperator{\plim}{plim}

\newcommand{\cN}{\mathcal{N}}
\newcommand{\cF}{\mathcal{F}}

\newcommand{\RR}{\mathbb{R}}
\newcommand{\NN}{\mathbb{N}}
\newcommand{\hb}{\hat{\beta}}
\newcommand{\dPois}{\mathrm{Pois}}
\newcommand{\dUnif}{\mathrm{Unif}}

\usepackage{mathtools}


\DeclarePairedDelimiter{\norm}{\lVert}{\rVert}
\DeclarePairedDelimiter{\abs}{\lvert}{\rvert}
\DeclarePairedDelimiter{\scalp}{\langle}{\rangle}
\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}




\begin{document}

%Madame de Pompadour%\footnote{official chief mistress of King Louis XV} 
%was born exactly 204 years earlier, on 29 December 1721. 
%Her original name was Jeanne Antoinette Poisson.





% \subsection*{Final demo :)}

%This is a demo version of the final exam. 
%Final will consist of 6 problems: generating functions in probability, Polya-Redfield counting, planarity and Euler formula, graph colouring, linearity of expected value and one old topic.
% midterm: two problems on binomial identities, two problems on recurrent equations, two problems on generating functions. 
%Some problems will be similar to demo 

% You may use one A4 cheat sheet in any form handwritten, printed, aquarelle painting\dots{ } 

\begin{enumerate}
    
    \item {[generating functions in probability]} 
    I throw a fair coin until the sequence $HHH$ appears. % The coin is not fair and shows $H$ with probability $0.4$.
\begin{enumerate}[label=\alph*)]
    \item Draw the directed graph for this game.
    \item Setup a system of equations for the sequences with given endpoint and starting point at the start of the game. 
    \item Explain how would you calculate the  expected number of \emph{tails} from the start till the end of the game.
    
    Note: you don't need to do actual calculations nor to solve the system, just describe how can you extract the expected value using derivatives. 
\end{enumerate}

    \item {[Polya-Redfield counting]}  Consider a $3\times 3$ mosaic of gems, where each gem may be of one of $k$ possible types.
    We consider mosaics identical if the can be transformed one into another by rotation. \emph{Flips are not available!} 
\begin{enumerate}[label=\alph*)]
    \item List all elements of the group of actions that act on mosaics.
    \item Write each action as a combination of cycles and provide its cyclic monomial. 
    \item Find the cyclic index and the number of different possible mosaics for $k=10$ gem types available.
\end{enumerate}
    \item {[planarity and Euler formula]} Rhombicosidodecahedron has $20$ triangular faces, $30$ square faces and some number of pentagonal faces. 
    The total number of vertices is $60$. 

    \begin{enumerate}[label=\alph*)]
        \item How many pentagonal faces does rhombicosidodecahedron have?
        \item How many edges are meeting at every vertex?
    \end{enumerate}

    \item {[graph colouring]} Consider the graph with three components. The first component is a $C_4$ graph, the second one — a tree with $10$ vertices,
    the third one — an isolated vertex. Components are non connected.
    
    \begin{enumerate}[label=\alph*)]
        \item Draw an example of such a graph.
        \item You have $10$ colours available. In how many ways can you color the vertices of this graph if adjacent vertices should be painted differently?
    \end{enumerate}

        
    \item {[linearity of expected value]} We put randomly $7$ pencils in $10$ boxes one by one choosing each box with equal probabilities.
\begin{enumerate}[label=\alph*)]
  \item What is the average number of empty boxes?
  \item What is the average number of boxes with more than one pencil?
\end{enumerate}


    \item {[old topic]} The generating sequence $g(t)$ of an integer sequence $(c_0, c_1, c_2, \dots)$ satisfies the equation 
    \[
    g(t) = 1 + t \cdot g(t) \cdot g(t).
    \]
    \begin{enumerate}
        \item Solve the equation for $g(t)$.
        \item Which of the two roots should we choose if all $c_n$ are positive?
        \item Find $c_3$.
    \end{enumerate}


\end{enumerate}


\end{document}

